---
title       : Teaching R
subtitle    : Part 1
author      : Low Yi Xiang
framework   : revealjs        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : dark      # 
revealjs    :
  theme: night
  center: "false"
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft,selfcontained}
knit        : slidify::knit2slides
---

## R Prog. Part 2

<br>

Done by: Low yi Xiang

<br>

[Linkedin Profile](https://www.linkedin.com/in/yi-xiang-low-b349137b)

<br>

[Jump To Contents](#/3)

---&vertical

## R DataFrames 

A data frame is used for storing data tables. It is a list of vectors of equal length. For example, the following variable df is a data frame containing three vectors n, s, b.

Example:

```{r}
name <- c("mary","john","tom")
grades <- c(89,60,74)
df_student_grades <- data.frame(name,grades)
print(df_student_grades)
```

This tutorial will teach you the basics of manuiplating csv (excel) files within R to perform analysis, joining data, etc.

This tutorial also assumes you have installed Rstudio and know how to install packages.

***
## Dataframes (Indexing)

Dataframes can be index by their row and columns using a matrix like structure. 

For example to extract the first row and first column value:

```{r}
df_student_grades[1,1]
```

If only the first entry (row) is of interest:
```{r}
df_student_grades[1,]
```

Or if only the grades are of interest:
```{r}
df_student_grades[,2] #the second column represent the grades 
```

***

## Dataframes (Indexing - Cont)

To extract the column names of a dataframe:

```{r}
names(df_student_grades)
```

We can also use column names instead of numerical value to extract a column. 
Note* this is also true for rows. 

```{r}
df_student_grades[,"grades"]
```

We can edit a specific cell in the dataframe:

```{r}
df_student_grades[3,"grades"] <- 80 #edit grades for the 3rd row 
df_student_grades[3,"grades"]
```

---

## Install Libraries required 

Please install the following libraries by running these codes:

```{r, eval = FALSE}
install.packages(c("dplyr","tidyr","datasets","ggplot2"))
```

<br>

Please check that you can install & run these libraries.
```{r, message=FALSE}
library(dplyr);
library(tidyr);
library(datasets);
library(ggplot2)
```

---

## Learning DataFrames

#### Reading/Writing Files 

- [Set Working Directory](#/4)

- [Reading/Writing Excel files, DataTypes](#/5)

#### Understanding DPLYR

- [Introduction / CheatSheet](#/6)

- [Filtering](#/7)

- [Select / Contains / Start/End](#/8)

- [Piping](#/9)

- [Introduction to Chaining](#/10)

- [Arrange](#/11)

- [Create New Columns](#/12)

- [Summarizing](#/13)

- [Grouped Operations](#/14)

- [Combine Datasets (Advanced)](#/15)

- [Window Functions (Advanced)](#/16)

- [Reshaping Data (Advanced)](#/17)

---

## Set Working Directory 

In R, you can set the working directory you wish to find the files. Please create a folder <q><b>named R_dataframe</q></b> in a location you prefer.  

Then, set your working directory as follows: (slightly different for different OS users)

```{r, eval = FALSE}
setwd('/Users/lowyix/Desktop/R_dataframe/')
```

Alternatively, you can nagivate to the help bar and click <q>Session - Set Working Directory - Choose directory </q> and then select the folder R_dataframe

To get your current working directory,

```{r, eval = FALSE}
getwd()
```

```{r, echo = FALSE}
#code to print out 
print("/Users/lowyix/Desktop/R_dataframe/")
```

To List all the files in the working directory, you can:
```{r,eval= FALSE}
list.files()
```

```{r, echo = FALSE}
#code to print out 
print(character(0))
```

[Back To Contents](#/3)

---&vertical

## Reading/Writing Excel Files 

lets load a dataset from one of the libraries; [diamonds](http://docs.ggplot2.org/0.9.3.1/diamonds.html)

```{r}
diamonds_df <- diamonds #if the code fails to work, check that you have loaded ggplot2 
```

#### Writing out dataframes 

To write out an excel (csv) file, 
```{r,eval=FALSE}
write.csv(diamonds_df,"diamonds_data.csv", row.names = FALSE)
list.files()
```

```{r, echo = FALSE}
#code to print out 
print("diamonds_data.csv")
```

More information can be found with:
```{r, eval=FALSE}
?write.csv
```

<br>

[Back To Contents](#/3)

***

## Reading in Dataframes

```{r,eval=FALSE}
diamonds_df <- read.csv("diamonds_data.csv")
head(diamonds_df) #more on this later 
```
```{r, echo = FALSE}
#code to print out 
diamonds_df <- diamonds
head(diamonds_df) #more on this later 
```
<br>

In R, you can manage your files too, such as deleting the files in the directory 
```{r,eval=FALSE}
file.remove("diamonds_data.csv")
```

<br>

[Back To Contents](#/3)


***

## Other Packages

There are other packages avaliable on CRAN to read/write excel tables and other formats such as xls or xlsx. 

However generally CSV is the prefered format as it is machine readable. (Remember that formats such as XLSX allows merge cells, multiple sheets per file etc)

<br>
<br>

Example of such packages are 

> - [openxlsx](https://cran.r-project.org/web/packages/openxlsx/openxlsx.pdf)
> - [readr](https://cran.r-project.org/web/packages/readr/readr.pdf)
> - [readxl](https://cran.r-project.org/web/packages/readxl/readxl.pdf)

<br>
<br>

[Back To Contents](#/3)

*** 

## Data Types 

In R Dataframes, there are mainly 3 types of columns, <u>Factors</u>, <u>Characeers</u> and <u>Integer/Numeric</u>. 

<br>

You can convert one class to another with commands like 
```{r,eval = FALSE}
as.integer(x)
as.numeric(x)
as.character(x)
as.factor(x)
```

<br>

[Back To Contents](#/3)

***

## Integer/Numeric 

Integer values means whole numbers only. For instance,

```{r}
as.integer(1.6)
```

<br>

Numeric values means "floats" or "decimals". You can convert "strings" or integers to numerics 

```{r}
string_one <- "1.6"
as.numeric(string_one)
as.integer(string_one)
```

<br>

[Back To Contents](#/3)

***

## Character

Character types are essentially "strings". For instance:

```{r}
as.character(c(1,2,3))
```

## Factors

Factors are essentially categories or "Groups" within the data. There are particularly useful when you want to do factor level operations - generally used when plotting graphs. 

They are also useful when you want to restrict operations within a Dataframe Factor Column.

```{r}
as.factor(c(1,2,3))
```

[Back To Contents](#/3)

***

## Factors (cont)
In the dataframe diamonds_df, the "str" command runs a quick summary and we can see the "cut","color" and "clarity" are factor columns.

```{r}
str(diamonds)
```

The different types of cut of a diamonds are:
```{r}
diamonds_df %>% select(cut) %>% distinct()
```

[Back To Contents](#/3)

***

## Factors (cont)

If one tries to edit / add data that is not within the 5 categories, an error will occur. Observe:

```{r}
diamonds_df[1,"cut"]  #this should return a value "Ideal"
```

<br>

Suppose the value is changed to another category within the list, such as Premium,
```{r}
diamonds_df[1,"cut"] <- "Premium" #this does not produce any error
```

The "NA" value will be generated instead:

```{r}
diamonds_df[1,"cut"]
```

<br>

[Back To Contents](#/3)

***

## Factor (cont)

If the value is not within the list:
```{r}
diamonds_df[1,"cut"] <- "testing" #this does not produce any error
diamonds_df[1,"cut"]
```

This is particularly useful when you want to make sure certain columns only accept certain values.

Lastly, reset the diamonds_df to its orginial value with the following code. 

```{r}
diamonds_df <- diamonds
```

<br>

[Back To Contents](#/3)

***

## Factors (Advance)

Just as an side note, it is possible to change the "levels" of the columns. 

The default settings in R is such that they will arrange the levels according to numeric/alphabetical order. Sometimes this may not be ideal (especially in plots), this can be corrected by the following steps:

```{r}
new_diamonds_df<- ggplot2::diamonds
levels(new_diamonds_df$cut)
```

<br>

In the case of diamonds, suppose we want the order Premium, Very Good, Good, Ideal, and Fair instead, we define a new numeric reflecting the previous levels shown above.

Note* - If you do not understand R indexing it is ok, the point of this slide is to show levels manipulation. 

```{r}
new_ordering <- c(4,3,2,5,1)
levels(new_diamonds_df$cut)[new_ordering] #observe the new changes 
```

<br>

[Back To Contents](#/3)

***

## Factors (Advance - Cont)

After defining the numeric vector, replace the orignial column with the new order:

<br>

```{r}
new_diamonds_df$cut <- factor(new_diamonds_df$cut,
                              levels = levels(new_diamonds_df$cut)[new_ordering])  

levels(new_diamonds_df$cut)
```

<br>

We can see that the levels are readjusted to the levels we desire! 

<br>

[Back To Contents](#/3)

---&vertical

## Introduction to DPLYR 

DPLYR is a package written by [Hadley Wickham](https://en.wikipedia.org/wiki/Hadley_Wickham) who is currently Chief Scientist at RStudio. 

It is a fast, consistent tool for working with data frame like objects, both in memory and out of memory.

> - [Github](https://cran.r-project.org/web/packages/dplyr/dplyr.pdf)

> - [Cheatsheet](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf)

Note* - The Cheatsheet covers a significant portion of DPLYR and a short description of each function. 

[Back To Contents](#/3)

---&vertical

## Filtering 

Lets look at the dataset again:

```{r}
head(diamonds_df)
```

As well as the names:
```{r}
names(diamonds_df)
```

<br>

[Back To Contents](#/3)

***

## Filtering (Cont)

Suppose we want to look at diamonds with color 'D':

```{r}
diamonds_df_color_D <- filter(diamonds_df,color == 'D')
head(diamonds_df_color_D,3)
```

Suppose we want to look at diamonds with a larger carat, say size 2 and above:
```{r}
diamonds_df_carat_abv2 <- filter(diamonds_df,carat >= 2)
head(diamonds_df_carat_abv2,3)
```

<br>

[Back To Contents](#/3)

***

## Filtering (Cont)

You can combine 'and' conditions or 'or' conditions as follows:

Example 'and':
```{r}
diamonds_df_color_D_and_carat_abv2 <- filter(diamonds_df, color == "D" & carat >= 2)
head(diamonds_df_color_D_and_carat_abv2,3)
```

<br>

Example 'or'
```{r}
diamonds_df_color_D_or_carat_abv2 <- filter(diamonds_df, color == "D" & carat >= 2)
head(diamonds_df_color_D_or_carat_abv2,3)
```

<br>

[Back To Contents](#/3)

---&vertical

## Selecting Columns

You can select Columns with the <u>"select"</u> command. For example:

```{r}
diamond_df_selected <- select(diamonds_df,carat,cut)
head(diamond_df_selected,3)
```

<br> 

There are other kind of commands that can be used in conjunction with Dplyr, for example <u>"contains"</u>:

```{r}
diamonds_df_contains_ri <- select(diamonds_df, contains("ri"))
head(diamonds_df_contains_ri,3)
```

<br>

[Back To Contents](#/3)

***

## Selecting Columns (Cont)

Example <u>'starts_with'</u>:
```{r}
diamonds_df_start_with_c <- select(diamonds_df, starts_with("c"))
head(diamonds_df_start_with_c,3)
```
Example <u>'ends_with'</u>:
```{r}
diamonds_df_ends_with_t <- select(diamonds_df, ends_with("t"))
head(diamonds_df_ends_with_t,3)
```

<br>

More information can be found in the [Cheatsheet](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf)

<br>

[Back To Contents](#/3)

---&vertical

## Piping

To understand piping, lets first understand a mock example by creating a function.

```{r}
multiply_each_other <- function(x,y){
  return(x*y)
}
```

Then, we can call this function as follows:
```{r}
multiply_each_other(2,4)
```

With Piping, we can perform a similar operation as follows:
```{r}
2 %>% multiply_each_other(4)
```

<br> 

[Back To Contents](#/3)

***

## Piping (Cont)

In other words, when you pipe an object to another function, it "fills" up the first argument of the new function. 

Observe the two codes we used earlier:
```{r}
df1 <- select(diamonds_df,carat)
head(df1,3)
```
vs
```{r}
df2 <- diamonds_df %>% select(carat)
head(df2,3)
```

<br>

[Back To Contents](#/3)

---&vertical

## Introduction to Chaining

The advantage of piping means it allows us to do chaining. To explain this concept, lets assume that you have to select some columns and filter by certain conditions. You would either:

```{r}
df_select <- select(diamonds_df, starts_with("c"))
df_select_filter <- filter(df_select, carat >=2)
head(df_select_filter,3)
```

OR:
```{r}
df_select_filter2 <- filter(select(diamonds_df, starts_with("c")),carat>=2)
head(df_select_filter2,3)
```

Notice the code can be a little hard to read - What if you have multiple functions to run?

[Back To Contents](#/3)

***

## Introduction to Chaining (Cont)

This is where Chaining is very useful by combining piping, 

Observe:
```{r}
df4 <- diamonds_df %>% 
  select(starts_with("c")) %>%
  filter(carat >= 2)

head(df4,3)
```

In the subsequent examples, you will see more of piping in action. 

Additional Note* - Piping exists in many other packages, such as tidyr, ggvis, leaflet and many other libraries.

[Back To Contents](#/3)

---&vertical

## Arranging

To sort the data by a particular column,

```{r}
diamonds_df %>% arrange(carat) %>% head(3)
```

You can arrange by two columns as well, for example by carat and clarity:
```{r}
diamonds_df %>% arrange(carat,clarity) %>% head(3)
```

[Back To Contents](#/3)

***

## Arranging (Cont)

To sort in a decreasing order, use the <u>"desc"</q></u> function:
```{r}
diamonds_df %>% arrange(carat,desc(clarity)) %>% head(3)
```

Another example:
```{r}
diamonds_df %>% arrange(desc(carat),desc(price)) %>% head(3)
```

[Back To Contents](#/3)

---&vertical

## Creating New Columns

Creating new columns can be done with the <u>"mutate"</u> or <u>"transmute"</u>" columns. The difference between transmute and mutate is transmute drops its orginial columns.

For example, lets compute the price per carat: 

```{r}
diamonds_df %>% mutate(price_per_carat = price/carat) %>% 
  select(starts_with('c'),price_per_carat) %>% head(3)
```

With transmute:
```{r}
diamonds_df %>% transmute(price_per_carat = price/carat) %>% head(3)
```

[Back To Contents](#/3)

***

## Creating New Columns (Cont)

You can combine ifelse / functions along with <u>mutate</u> functions.

For example you want to create a new column to differentiate large diamonds (abv 1 carat) and small diamonds. 

```{r}
set.seed(5) #set seed allows everyone to have the same results
diamonds_df %>% 
  mutate(big_or_small = ifelse(carat >=1, "big","small")) %>% 
  select(starts_with('c'),big_or_small) %>%
  sample_n(5) %>%
  head(5)
```

<br>

[Back To Contents](#/3)

***

## Creating New Columns (Cont)

You can use functions as well:
```{r}
carat_big_small <- function(x){
  temp = ifelse(x>=1,"big","small")
  return(temp)
}

set.seed(5) #set seed allows everyone to have the same results
diamonds_df %>% 
  mutate(big_or_small = carat_big_small(carat)) %>% 
  select(starts_with('c'),big_or_small) %>%
  sample_n(5) %>%
  head(5)

```

<br>

[Back To Contents](#/3)

---&vertical

## Summarizing

There are numerous summarize functions, lets look at some examples:

```{r}
diamonds_df %>% summarise(mean(table))
```

You can include multiple functions as follows:

```{r}
diamonds_df %>% summarise(mean = mean(table),
                          median = median(table),
                          unique_carat = n_distinct(carat),
                          max_carat = max(carat))
```

[Back To Contents](#/3)

---&vertical

## Grouped Data

<u>group_by</u> should be a familar operation for those who use SQL, let's see an example first:

We want to know the relationship between the cut and the price. 
```{r}
diamonds_df %>% 
  group_by(cut) %>% 
  summarise(mean(price))
```

<br>

[Back To Contents](#/3)

***

## Grouped Data (Cont)

What about the color?

```{r}
diamonds_df %>%
  group_by(color) %>%
  summarise(mean(price))
```

<br>

[Back To Contents](#/3)

***

## Grouped Data (Cont)

Suppose you want to find out the max price within each group

```{r}
diamonds_df %>%
  group_by(color, cut) %>%
  summarise(max_price = max(price)) %>%
  arrange(max_price) %>%
  head(10)
```

We can see that (E, Very Good) has a higher value than (D, Fair). This is because when using the group_by operation, it finds out the max price within each group. 

<br>

[Back To Contents](#/3)

***

## Grouped Data (Cont)

Sometimes you will need to <u>"ungroup()"</u> your data to remove the grouping, lets see an example:

Question : find out the highest average price within each group and then list the top 5 groups with the highest everage price. 

```{r}
df5 <- diamonds_df %>%
  group_by(color, cut) %>%
  summarise(mean_price = mean(price)) %>%
  arrange(desc(mean_price))

df5 %>% head(6)
```

We can see that (E,fair) has a higher price than (D,Ideal).

<br>


[Back To Contents](#/3)

***

## Grouped Data (Cont)

We can fix this with the <u>"ungroup"</u> option: 

```{r}
diamonds_df %>%
  group_by(color, cut) %>%
  summarise(mean_price = mean(price)) %>%
  ungroup() %>%
  arrange(desc(mean_price)) %>%
  head(10)
```

Now the highest 10 mean prices belong to J, I, H cuts. One can conclude that color affects the price more than the cut. 

<br>

Question:Find out which cut has the highest price within each color. 

[Back To Contents](#/3)

***

## Grouped Data (Cont)

Answer:

```{r}
diamonds_df %>% 
  group_by(color,cut) %>%
  summarise(mean_price = mean(price), count = n()) %>%
  ungroup() %>%
  group_by(color)%>%
  filter(mean_price == max(mean_price))
```

---&vertical

## Combine Datasets (Advance)

[Back To Contents](#/3)

There are two ways to join data set, either by joining, or binding. For joining data there is <u>left_join</u>, <u>right_join</u>, <u>inner_join</u>, <u>full_join</u>, <u>semi_join</u>, <u>anti_join</u>.

Define the following variables:
```{r}
data_A <- data.frame(names = c("amber","bryan","diana"), grades = c(84,93,56))
data_B <- data.frame(names = c("amber","bryan","charlie"), gender = c("F","M","M"))
```

left_join : join matching rows from B to A
```{r}
data_A %>% left_join(data_B, by = "names")
```

<br>

[Back To Contents](#/3)

***

## Combine Datasets (Adv-Cont)

Recall that dealing with factor with differnet levels will result in a warning message. To overcome this, lets change them to character levels:

```{r}
data_A <- data_A %>% mutate(names = as.character(names))
data_B <- data_B %>% mutate(names = as.character(names))
```

<br>

[Back To Contents](#/3)

***

## Combine Datasets (Adv-Cont)

left_join: join matching rows from b to a
```{r}
data_A %>% left_join(data_B, by = "names")
```
right_join:Join matching rows from a to b.
```{r}
data_A %>% right_join(data_B, by = "names")
```
inner_join:Join data. Retain only rows in both sets.
```{r}
data_A %>% inner_join(data_B, by = "names")
```

<br>

[Back To Contents](#/3)

***

## Combine Datasets (Adv-Cont)

full_join: Join data. Retain all values, all rows.
```{r}
data_A %>% full_join(data_B, by = "names")
```

semi_join: All rows in a that have a match in b.
```{r}
data_A %>% semi_join(data_B, by = "names")
```
anti_join: All rows in a that do not have a match in b.
```{r}
data_A %>% anti_join(data_B, by = "names")
```

<br>

[Back To Contents](#/3)

***

## Combine Datasets (Adv-Cont)

You can bind datasets together as well:

bind_rows:
```{r}
data_C <- data.frame(names = c("cheryl","elieen","faith"), grades = c(40,60,100)) %>%
  mutate(names = as.character(names))

data_A %>% bind_rows(data_C)
```

bind_cols:
```{r}
data_A %>% bind_cols(data_C)
```

[Back To Contents](#/3)

---&vertical

## Window Functions (Advanced)

Window Functions are essentially functions that performs specific analysis. 

more details can be found [here](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf)

Suppose you want to find your cumulative sales each month: 

```{r}
data_sales <- data.frame(month = 1:12, sales = rnorm(12,10000,2000))
data_sales %>% mutate(cumulative_sales = cumsum(sales))
```

[Back To Contents](#/3)

---&vertical

## Reshaping Data (Advanced)

There are two types of data, "wide" and "long"/"narrow" format. 

The following shows a "wide" format.

<br>

```{r kable, echo = FALSE}
data_D <- data.frame(names = c("Amber","Bryan","Charlie"), 
                     weight = c(60,90,80),
                     height = c(152,184,178)) %>%
  mutate(names = as.character(names))
library(knitr)
kable(data_D)
```

<br>

[Back To Contents](#/3)

***
## Reshaping Data (Adv-Cont)

The following shows a 'long' format:

<br>

```{r kable2, echo = FALSE}
data_D_2 <- data_D %>% gather(key = variable, value = measurements, weight,height)
library(knitr)
kable(data_D_2)
```

<br>

More information can be found [here](https://en.wikipedia.org/wiki/Wide_and_narrow_data). There are many uses cases where you need wide over long data. 

<br>

For examples, refer [here](http://www.theanalysisfactor.com/wide-and-long-data/)

[Back To Contents](#/3)

***

## Reshaping Data (Adv-Cont)

Lets look at a more complicated example with the following mock dataset of 2 students over 2 years:
```{r, echo = FALSE}
data_E <- data.frame(names=c("amber","bryan"), 
                     grades_Q1_2015 = c(60,80),
                     grades_Q2_2015 = c(54,83),
                     grades_Q3_2015 = c(90,74),
                     grades_Q4_2015 = c(87,76),
                     grades_Q1_2016 = c(60,79),
                     grades_Q2_2016 = c(58,87),
                     grades_Q3_2016 = c(40,85),
                     grades_Q4_2016 = c(93,81)
                     )
print(data_E)
```

Suppose i want to find out the average score of each year, i have to write the following code:
```{r}
data_E %>% transmute(Y2015 = 0.25*(grades_Q1_2015+grades_Q2_2015+grades_Q3_2015+grades_Q4_2015),
                     Y2016 = 0.25*(grades_Q1_2016+grades_Q2_2016+grades_Q3_2016+grades_Q4_2016))
```

This requires lots of coding + sometimes it may not be just 8 columns in this example ; what if the data for sales is over a 10 years period and a monthly breakdown is required?

<br>

[Back To Contents](#/3)

***

## Reshaping Data (Adv-Cont)



This is when converting wide to long data can be extremely helpful for further analysis. Usually as a rule of thumb, long format is for analysis while wide format is for viewing consumption (i.e on a presentation).

For the same analysis in the previous slide: 

```{r}
data_E %>%
  gather(key = variable, value = score, 
         grades_Q1_2015,grades_Q2_2015,grades_Q3_2015,grades_Q4_2015,
         grades_Q1_2016,grades_Q2_2016,grades_Q3_2016,grades_Q4_2016) %>%
  separate(variable, c("grades","quarter","year"), sep = "_") %>% #More information in next slides
  group_by(names,year) %>%
  summarise(mean(score))
```

[Back To Contents](#/3)

***

## Reshaping Data (Adv-Cont)

Lets explain the first part how <u>"gather"</u> works

```{r}
data_E %>%
  gather(key = variable, value = score, 
         grades_Q1_2015,grades_Q2_2015,grades_Q3_2015,grades_Q4_2015,
         grades_Q1_2016,grades_Q2_2016,grades_Q3_2016,grades_Q4_2016) %>%
  head(2)
```

OR:

```{r}
data_E %>%
  gather(key = variable, value = score, 
         -names) %>% #notice the hypen sign
  head(2)
```

In other words, you specify the new key column name, the new value column name, followed by either the key columns (with the minus sign) or the value columns. 

[Back To Contents](#/3)

***

## Reshaping Data (Adv-Cont)

Heres another example with the iris dataset which is found in the documentation of gather:

By specifying the "value" columns:
```{r}
mini_iris <- iris[c(1, 51, 101), ]
gather(mini_iris, key = flower_att, value = measurement,
       Sepal.Length, Sepal.Width, Petal.Length, Petal.Width)
```


[Back To Contents](#/3)

***

## Reshaping Data (Adv-Cont)

By specifying the "key" columns:
```{r}
gather(mini_iris, key = flower_att, value = measurement, -Species)
```


[Back To Contents](#/3)

***

## Reshaping Data (Adv-Cont)

Usually the column names in a database or an excel sheet is uniformly labeled, such as the example provided:

```{r}
names(data_E)
```

After the gather step, we arrive at and we want to split the names to 3 different columns:

```{r}
data_E %>%
  gather(key = variable, value = score, 
         -names) %>% #notice the hypen sign
  head(2)
```

<br>

[Back To Contents](#/3)

***

## Reshaping Data (Adv-Cont)

We can achieve this by the <u>"separate"</u> function: 

The way to use it is separate(col_name,new_col_names, separator)
```{r}
data_F <- data_E %>%
  gather(key = variable, value = score, 
         -names) %>% #notice the hypen sign
    separate(variable, c("grades","quarter","year"), sep = "_") 
data_F %>% head(2)
```

you can type ?separate in your console to find out more! the opposite of this is unite. 

<br>

[Back To Contents](#/3)

***

## Reshaping Data (Adv-Cont)

Example <u>unite</u>:

unite(new_col_name,col_to_merge(s),separator)
```{r}
data_G <- data_F %>% unite(variable,grades,quarter,year,sep="_") 
data_G %>% head(5)
```

you can type ?unite in your console to find out more! 

<br>

[Back To Contents](#/3)

***

## Reshaping Data (Adv-Cont)

The opposite of gather is <u>spread</u>:
gather(variable_column,value_column)
```{r}
data_G %>% spread(key = variable, value = score)
```

<br>

[Back To Contents](#/3)

---&vertical

# End!

<br>
<br>
<br>
[Back To Contents](#/3)









